---
title: "STAR+WASP reduces reference bias in the allele-specific mapping of RNA-seq reads: Manuscript Figures"
author: "Rebecca Asiimwe & Alexander Dobin"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    downcute_theme: "chaos"
editor_options: 
  chunk_output_type: console
---


## Loading Required Libraries
```{r, warning=FALSE, message=FALSE}
library(splitstackshape)
library(dplyr)
library(ggplot2)
library(VennDiagram)
library(RColorBrewer)
library(gridExtra)
library(cowplot)
library(knitr)
library(kableExtra)
library(tidyverse)
library(Hmisc)
library(ggridges)
library(venn)
library(ggthemes)
library(stringr)
library(splitstackshape)
library(chron)
library(magicfor)
library(ggtext)
library(reshape2)
library(plyr)
library(formattable)
library(data.table)
library(ggrepel)
require(MASS)
require(scales) 
```

```{r}
setwd("~/projects/run_env/alpha_star_wasp_benchmarking/downstream_analysis/Alpha/Data/")
```

## Loading data extracted from alignments and subsequent data calls
```{r, warning=FALSE, message=FALSE}
num_reads_reshaped <- read.csv("num_reads_reshaped.csv")
stats_vW_Tags <- read.csv("stats_vW_Tags.csv")
df_combined <- read.csv("df_combined.csv")
wall_clock_melted_converted <- read.csv("wall_clock_melted_converted.csv")
benchmark_rez_melted <- read.csv("benchmark_rez_melted.csv")
```

## Reads Overlapping Variants and nReads per Sample
```{r, fig.heigt=10, fig.width=7, warning=FALSE, message=FALSE}
STAR_subset <- benchmark_rez_melted %>% filter(Run == "STAR")
df_orders <-
  STAR_subset[order(
    STAR_subset$Sample,
    STAR_subset$Average_input_read_length,
    STAR_subset$Number_of_input_reads
  ), ]
order_df <- as.character(rev(unique(df_orders$Sample)))

df_combined$state1 <-
  ordered(
    df_combined$state1 ,
    levels = c(
      "All Alignments",
      "Alignment Failed WASP Filtering",
      "Alignment Passed WASP Filtering"
    )
  )

global_colors <-  c("darkorange3", "dodgerblue4", "firebrick4")

## Adding read lengths
read_lengths <- benchmark_rez_melted[, c(2, 7)]
read_lengths <- read_lengths[!duplicated(read_lengths$Sample),]
num_reads_reshaped_nreads <-
  inner_join(num_reads_reshaped, read_lengths, by = "Sample")
num_reads_reshaped_nreads$Number_of_input_reads_mil <-
  (num_reads_reshaped_nreads$Number_of_input_reads) / 1000000

(b <-
    num_reads_reshaped_nreads[order(num_reads_reshaped_nreads$Flag, decreasing = T), ] %>%  filter(Flag == "vW_Tagged Reads")  %>%
    ggplot() +
    geom_bar(
      aes(
        x = factor(Sample, levels = order_df),
        y = perc,
        group = Sample,
        fill = factor(Flag, levels = c("vW_Tagged Reads", "Reads with no Tag"))
      ),
      stat = "identity",
      width = 0.99,
      alpha = 0.9,
      color = "white"
    ) +
    geom_text(
      aes(
        x = Sample,
        y = perc,
        label = paste0(round(perc, 1), "%")
      ),
      size = 4,
      position = position_stack(vjust = 0.5)
    ) +
    geom_text(
      aes(
        x = Sample,
        y = perc,
        label = round(Number_of_input_reads_mil, 1)
      ),
      size = 4,
      position = position_stack(vjust = 1.016),
      color = "gray40"
    ) + 
    theme_test(base_size = 13) +
    scale_color_manual(values = c("gray60", "gray60")) +
    scale_fill_manual(values = "gray70") +
    labs(y = "Reads Overlapping Variants (% of Total)", x = "") +
    theme(legend.title = element_blank(), legend.position = "none") +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 10)) + scale_x_discrete(expand = c(0, 0))
) + theme(
  axis.title.y = element_text(size = 12),
  axis.text.x = element_text(
    size = 12,
    angle = 90,
    hjust = 1
  ),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 7)
)          
```

## Classification of WASP-Rejected Reads
```{r, warning=FALSE, message=FALSE}
stats_vW_Tags$vW_Tag_desc <-
  as.character(stats_vW_Tags$vW_Tag_desc)
stats_vW_Tags$vW_Tag_desc[stats_vW_Tags$vW_Tag_desc == "Variant Base in Read is N/non-ACGT"] <- "Variant Base in Read is N"

(d <-
    stats_vW_Tags %>% filter(vW_Tag != 1) %>% filter(round != 0) %>%
    ggplot() +
    geom_bar(
      aes(
        x = factor(Sample, levels = order_df),
        round(Freq, 3),
        group = Sample,
        fill = vW_Tag_desc
      ),
      stat = "identity",
      width = 0.99,
      alpha = 0.8
    ) +
    geom_text(
      aes(
        x = Sample,
        y = Freq,
        label = paste0(round(Freq, 2), "%")
      ),
      size = 4,
      position = position_stack(vjust = 0.5)
    ) +
    scale_fill_manual(
      values = c(
        "Variant Base in Read is N" = "gray15",
        "Remapped Read did not Map" = "tan2",
        "Remapped Read Maps to Different Locus" =
          "dodgerblue4",
        "Read Overlaps too Many Variants" = "firebrick4"
      )
    ) +
    theme_test(base_size = 13) +
    labs(y = "WASP-Rejected Reads (%)", x = "") +
    theme(axis.text.x = element_text(
      angle = 0,
      hjust = 1,
      vjust = 0.5,
      size = 10
    )) +
    scale_y_continuous(expand = c(0.03, 0), limits = c(0, 4.5)) +
    scale_x_discrete(expand = c(0, 0)) + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +   theme(axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 10
    )) +
    theme(
      legend.title = element_blank(),
      legend.position = c(0.22, 0.8)
    ) +
    guides(fill = guide_legend(nrow = 4, byrow = FALSE))
) + theme(
  axis.title.y = element_text(size = 12),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  legend.text = element_text(size = 9)
)
```

## Reference Bias - Three-Way Comp
```{r, fig.height=5, fig.width=4, warning=FALSE, message=FALSE}
scale_y <- scales::trans_new(
  "signed_log",
  transform = function(x)
    sign(x) * log1p(abs(x)),
  inverse = function(x)
    sign(x) * expm1(abs(x))
)

df_combined <- read.csv("df_combined.csv")

df_combined$state1 <-
  ordered(
    df_combined$state1 ,
    levels = c(
      "All Alignments",
      "Alignment Failed WASP Filtering",
      "Alignment Passed WASP Filtering"
    )
  )

df_combined$state2 <- df_combined$state1
df_combined$state2 <-
  as.character(df_combined$state2)
df_combined$state2[df_combined$state2 == "All Alignments"] <-
  "All Alignments"
df_combined$state2[df_combined$state2 == "Alignment Passed WASP Filtering"] <-
  "Alignment Passed WASP Filtering"
df_combined$state2[df_combined$state2 == "Alignment Failed WASP Filtering"] <-
  "Alignment Failed WASP Filtering"


df_combined$state2 <-
  ordered(
    df_combined$state2 ,
    levels = c(
      "All Alignments",
      "Alignment Failed WASP Filtering",
      "Alignment Passed WASP Filtering"
    )
  )
unique(df_combined$state2)

(c <- df_combined %>%
    ggplot(aes(
      x = state2,
      y = diff,
      fill = state2,
      color = state2
    )) +
    geom_boxplot(alpha = 0.9, size = 0.2) + geom_point(size = 1) +
    labs(y = paste0("Reference Bias (REF% - ALT%)"), x = "") +
    geom_vline(xintercept = 0) +
    geom_hline(
      yintercept = 0,
      color = "gray30",
      linetype = "dashed",
      size = 0.3
    ) +
    scale_color_manual(values = c("gray40", "firebrick4", "steelblue4"))  +
    scale_fill_manual(values = c("gray40", "firebrick4", "steelblue4"))  +
    theme_light(base_size = 12) +
    scale_y_continuous(
      trans = scale_y,
      breaks = c(-5, 0, 5, 10, 15, 20, 50, 60, 70, 80, 90, 100),
      limits = c(-5, 100),
      labels = c(-5, "0", "5", "10", "", "", "50", "", "70", "", "", "100")
      
    ) +
    annotation_logticks(
      sides = "l",
      outside = F,
      short = unit(0.05, "cm"),
      mid = unit(0.05, "cm"),
      long = unit(0.2, "cm")
    ) +
    
    theme(legend.position = "right", legend.title = element_blank()) +
    theme(
      axis.title.y = element_text(size = 12),
      axis.text.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.ticks.x = element_blank(),
      legend.text = element_text(size = 9)
    ) + theme(
      legend.title = element_blank(),
      legend.position = c(0.4, 0.15),
      legend.background = element_rect(fill = alpha('white', 0))
    )
)
```

## Reference Bias - Two-Way Comp
```{r, fig.height=5, fig.width=4, warning=FALSE, message=FALSE}
(f <-
    df_combined %>% filter(
      state1 == "All Alignments"  |
        state1 == "Alignment Passed WASP Filtering"
    ) %>%
    ggplot(aes(
      x = state1, y = diff, fill = state1
    ))  +
    geom_boxplot(alpha = 0.9, size = 0.3) +
    geom_point(size = 1, color = "gray20") +
    labs(
      y = paste0("Reference Bias", "\n", "(REF% - ALT%)"),
      x = ""
    ) +
    scale_color_manual(values = c("firebrick4", "steelblue4"))  +
    scale_fill_manual(values = c("firebrick4", "steelblue4"))  +
    geom_hline(
      yintercept = 0,
      color = "gray20",
      linetype = "dashed",
      size = 0.4
    ) +
    theme_test(base_size = 13) +
    scale_y_continuous(limits = c(-5, 10)) +
    theme(
      legend.position = c(0.45, 0.15),
      legend.title = element_blank()
    ) +
    scale_x_discrete(
      labels = function(x)
        str_wrap(x, width = 12)
    ) + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  
) + theme(axis.title.y = element_text(size =12), axis.text.x = element_text(size =8), axis.text.y = element_text(size =8), legend.text = element_text(size =7))
```

## Mapping speed for STAR, WASP and STAR+WASP 
```{r, fig.height=5, fig.width=4, warning=FALSE, message=FALSE}
(e <- wall_clock_melted_converted %>% filter(Threads == "8 Threads") %>% 
    ggplot(aes(x = factor(Sample, levels=order_df), y = reads_per_hour/1000000, group=Run)) +
    geom_point(aes(color=factor(Run),shape=factor(Run),fill=factor(Run)), size=3)+
    scale_shape_manual(values=c(3, 17, 16))+ 
    scale_color_manual(values=global_colors)+   
    labs(y = "Million Reads/Hour", x="")+
    theme_light(base_size = 12) + theme(legend.title = element_blank(), legend.position = c(0.76,0.15),
                                        legend.background = element_rect(fill=alpha('white', 0)))+ 
    theme(strip.background =element_rect(fill="white", colour = "white"))+
    theme(strip.text = element_text(colour = 'black'), strip.text.x =  element_markdown(hjust = 0.5, size=1)) +
    theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0.5, size=10)) + 
    scale_y_log10(breaks=c(0,10,20,30,40,50,60,70,80,90, 100,200,300,400,500,600,700,800,900,1000), limits=c(10, 1000), labels=c("","10","","","","","","","","","100","","","","","","","","","1000"))+
    annotation_logticks(sides = "l", outside = F, short = unit(0.05, "cm"),mid = unit(0.05, "cm"),long = unit(0.2, "cm"))+
    coord_cartesian(clip = "off")) + theme(axis.title.y = element_text(size =12), axis.text.x = element_text(size =8), axis.text.y = element_text(size =8), legend.text = element_text(size =7))
```


## Figure 1 Panel (Main)
```{r, fig.height=10, fig.width=12, warning=FALSE, message=FALSE}
b <- b + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

(main <-
    plot_grid(
      b,
      c,
      d,
      e,
      align = c("v", "h"),
      rel_widths = c(2.1, 1.1),
      rel_heights = c(1, 1.35),
      labels = c("(b)", "(d)", "(c)", "(e)")
    ))

ggsave(file="manuscript_figures/Figure1.pdf", main,  width=12, height=9, device="pdf", dpi=700 )
ggsave(file="manuscript_figures/Figure1.pdf", main,  width=12, height=9, device="pdf", dpi=700 )
```

## Supplementary and Other Figures

## Number of Reads per Sample
```{r, fig.width=6, warning=FALSE, message=FALSE}
intro_order <- c(
    "Total_R1",
    "Total_R2",
    "PolyA_R1",
    "PolyA_R2",
    "Nucleus_PolyA_R1",
    "Nucleus_PolyA_R2",
    "Nucleus_nonPolyA_R1",
    "Nucleus_nonPolyA_R2",
    "HG00733",
    "NA19239",
    "HG00732",
    "HG00512",
    "NA19238",
    "NA19240",
    "HG00513",
    "HG00731"
  )

num_input_reads_per_sample <- read.csv("num_input_reads_per_sample_supp.csv")
num_input_reads_per_sample %>%
  ggplot(aes(
    reorder(
      x = factor(Sample, levels = intro_order),
      -Number_of_input_reads_initial / 1000000
    ),
    y = Number_of_input_reads_initial / 1000000
  )) +
  geom_bar(stat = "identity", fill = "gray65") +
  theme_test(base_size = 12) +
  scale_y_continuous(expand = c(0, 0), limits = c(
    0,
    max(num_input_reads_per_sample$Number_of_input_reads_initial) / 1000000 +
      1
  )) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(y = "Number of Input Reads (Million)", x = "") +
  theme(axis.text.x = element_text(
    angle = 90,
    hjust = 1,
    vjust = 0.5,
    size = 10
  ))
```


## Reference bias for all reads vs. reads that pass filtering for each sample
```{r, warning=FALSE, message=FALSE}
df_combined <- read.csv("df_combined.csv")

df_combined_sub <-
  df_combined %>% filter(state1 == "All Alignments" |
                                     state1 == "Alignment Passed WASP Filtering")
unique(df_combined_sub$state1)

df_combined$state1 <-
  ordered(
    df_combined$state1 ,
    levels = c("All Alignments",  "Alignment Passed WASP Filtering")
  )

df_combined_sub <- df_combined_sub[, c(2, 5, 7)]
df_combined_sub_reshaped <-
  reshape(
    df_combined_sub,
    idvar = "Sample",
    timevar = "state1",
    direction = "wide"
  )

colnames(df_combined_sub_reshaped) <-
  c("Sample", "REF-ALT After Filtering", "REF-ALT Unfiltered")

df_combined_sub_reshaped %>%
  ggplot(aes(x = `REF-ALT Unfiltered`, y = `REF-ALT After Filtering`,  color =
               Sample)) + geom_point(size = 2.5) + scale_color_manual(
                 values = c(
                   "#67000D",
                   "#A50F15",
                   "#CB181D",
                   "#EF3B2C",
                   "#A63603",
                   "#FD8D3C",
                   "#FDD0A2",
                   "#FDAE6B",
                   "#08519C",
                   "#2171B5",
                   "#4292C6",
                   "#6BAED6",
                   "#006D2C",
                   "#238B45",
                   "#41AB5D",
                   "#74C476"
                 )
               ) + scale_x_continuous(limits = c(0, 10)) + scale_y_continuous(limits = c(0, 10)) + theme_cowplot()
```


## Mapping speed for STAR, WASP and STAR+WASP - 8 threads
```{r, fig.height=5, fig.width=4, warning=FALSE, message=FALSE}
wall_clock_melted_converted <- read.csv("wall_clock_melted_converted.csv")

global_colors <-  c("darkorange3", "dodgerblue4", "firebrick4")

#Million Reads/Hour 8threads
(threads8 <-
    wall_clock_melted_converted %>% filter(Threads == "16 Threads") %>%
    ggplot(aes(
      x = factor(Sample, levels = order_df),
      y = reads_per_hour / 1000000,
      group = Run
    )) +
    geom_point(aes(
      color = factor(Run),
      shape = factor(Run),
      fill = factor(Run)
    ), size = 3) +
    scale_shape_manual(values = c(3, 17, 16)) +
    scale_color_manual(values = global_colors) +
    labs(y = "Million Reads/Hour", x = "") +
    theme_light(base_size = 12) + theme(legend.title = element_blank(), legend.position = "none") +
    theme(strip.background = element_rect(
      fill = "white", colour = "white"
    )) +
    theme(
      strip.text = element_text(colour = 'black'),
      strip.text.x =  element_markdown(hjust = 0.5, size = 12)
    ) +
    theme(axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 10
    )) +
    scale_y_log10(
      breaks = c(
        0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 2000
      ),
      limits = c(10, 2000),
      labels = c(
        "", "10", "", "", "", "", "", "", "", "", "100", "", "", "", "", "", "", "", "", "1000", 2000
      )
    ) +
    annotation_logticks(
      sides = "l",
      outside = F,
      short = unit(0.05, "cm"),
      mid = unit(0.05, "cm"),
      long = unit(0.2, "cm")
    ) +
    coord_cartesian(clip = "off")
)
```

## Mapping speed for STAR, WASP and STAR+WASP - 32 threads
```{r, fig.height=5, fig.width=5.5, warning=FALSE, message=FALSE}
(threads32 <-
    wall_clock_melted_converted %>% filter(Threads == "32 Threads") %>%
    ggplot(aes(
      x = factor(Sample, levels = order_df),
      y = reads_per_hour / 1000000,
      group = Run
    )) +
    geom_point(aes(
      color = factor(Run),
      shape = factor(Run),
      fill = factor(Run)
    ), size = 3) +
    scale_shape_manual(values = c(3, 17, 16)) +
    scale_color_manual(values = global_colors) +
    labs(y = "", x = "") +
    theme_light(base_size = 12) + theme(legend.title = element_blank(), legend.position = "right") +
    theme(strip.background = element_rect(
      fill = "white", colour = "white"
    )) +
    theme(
      strip.text = element_text(colour = 'black'),
      strip.text.x =  element_markdown(hjust = 0.5, size = 12)
    ) +
    theme(axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 10
    )) +
    theme(axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 10
    )) +
    scale_y_log10(
      breaks = c(
        0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 2000
      ),
      limits = c(10, 2000),
      labels = c(
        "", "10",
        "", "", "", "", "", "", "", "", "100", "", "", "", "", "", "", "", "", "1000", 2000
      )
    ) +
    annotation_logticks(
      sides = "l",
      outside = F,
      short = unit(0.05, "cm"),
      mid = unit(0.05, "cm"),
      long = unit(0.2, "cm")
    ) +
    coord_cartesian(clip = "off")
)
```

```{r, fig.height=4.5, fig.width=8, warning=FALSE, message=FALSE}
plot_grid(
  threads8,
  threads32,
  nrow = 1,
  ncol = 2,
  labels = c("(a)", "(b)"),
  rel_widths = c(0.68, 0.99),
  hjust = 0.01,
  align = 'hv'
) 
```

## Distribution of Reads Overlapping Variants (/not)
```{r, fig.height=4.5, fig.width=9, warning=FALSE, message=FALSE}
num_reads <- as.data.frame(read.table("read_nums.txt"))

colnames(num_reads)
colnames(num_reads) <- c("Number_of_Reads", "Read_File_Path")
num_reads <-
  cSplit(num_reads, "Read_File_Path", "/", direction = "wide")
num_reads <- num_reads[, -c(2:9, 11)]
num_reads$Read_File_Path_09 <-
  as.character(num_reads$Read_File_Path_09)
num_reads$Read_File_Path_11 <-
  as.character(num_reads$Read_File_Path_11)

colnames(num_reads) <- c("Number_of_Reads", "Sample", "Flag")
num_reads$Flag[num_reads$Flag == "WASP_Reads_Sorted_Unique"] <-
  "Num_Reads"
num_reads$Flag[num_reads$Flag == "STAR_WASP_vW_Tagged_Reads_Unique"] <-
  "vW_Tagged_Reads"

#Converting data frame to short format
num_reads_reshaped <-
  reshape(num_reads,
          idvar = "Sample",
          timevar = "Flag",
          direction = "wide")
colnames(num_reads_reshaped)[2] <- "Num_Reads"
colnames(num_reads_reshaped)[3] <- "vW_Tagged_Reads"

# Mutating frequencies
num_reads_reshaped <-
  num_reads_reshaped %>% mutate("perc_tagged" = ((vW_Tagged_Reads / Num_Reads) *
                                                   100))
num_reads_reshaped <-
  num_reads_reshaped %>% mutate("perc_untagged" = 100 - perc_tagged)

# Plotting distributions
num_reads_reshaped_tagged <-
  as.data.frame(num_reads_reshaped[, c(1, 4)])
num_reads_reshaped_notags <-
  as.data.frame(num_reads_reshaped[, c(1, 5)])
colnames(num_reads_reshaped_tagged)[2] <- "perc"
colnames(num_reads_reshaped_notags)[2] <- "perc"
num_reads_reshaped_tagged$Flag <- "vW_Tagged Reads"
num_reads_reshaped_notags$Flag <- "Reads with no Tag"

num_reads_reshaped <-
  rbind(num_reads_reshaped_tagged, num_reads_reshaped_notags)


num_reads_reshaped$Flag <-
  ordered(num_reads_reshaped$Flag ,
          levels = c("vW_Tagged Reads", "Reads with no Tag"))

unique(num_reads_reshaped$Sample)
num_reads_reshaped <-
  num_reads_reshaped %>% filter(
    Sample != "NA12878_RAMPAGE_Rep" &
      Sample != "NA12878_RAMPAGE" & Sample != "NA12878_Small"
  )

## Cleaning up Sample names
unique(num_reads_reshaped$Sample)
#Removing NA12878 from all NA12878 derived samples
num_reads_reshaped$Sample <-
  gsub("NA12878_", "", num_reads_reshaped$Sample)
num_reads_reshaped$Sample <-
  gsub("Rep", "R2", num_reads_reshaped$Sample)
num_reads_reshaped$Sample[num_reads_reshaped$Sample == "Total"] <-
  "Total_R1"
num_reads_reshaped$Sample[num_reads_reshaped$Sample == "PolyA"] <-
  "PolyA_R1"
num_reads_reshaped$Sample[num_reads_reshaped$Sample == "Nucleus_PolyA"] <-
  "Nucleus_PolyA_R1"
num_reads_reshaped$Sample[num_reads_reshaped$Sample == "Nucleus_nonPolyA"] <-
  "Nucleus_nonPolyA_R1"

# Manuscript Plot Read_Distribution
num_reads_reshaped_plot <- num_reads_reshaped
num_reads_reshaped_plot$Flag <-
  as.character(num_reads_reshaped_plot$Flag)
num_reads_reshaped_plot$Flag[num_reads_reshaped_plot$Flag == "vW_Tagged Reads"] <-
  "Reads Overlapping Variants"
num_reads_reshaped_plot$Flag[num_reads_reshaped_plot$Flag == "Reads with no Tag"] <-
  "Reads not Overlapping Variants"
num_reads_reshaped_plot$Flag <-
  as.factor(num_reads_reshaped_plot$Flag)
num_reads_reshaped_plot[order(num_reads_reshaped_plot$Flag, decreasing = F), ] %>%
  ggplot() +
  geom_bar(
    aes(
      x = factor(Sample, levels = order_df),
      y = perc,
      group = Sample,
      fill = factor(
        Flag,
        levels = c("Reads Overlapping Variants", "Reads not Overlapping Variants")
      )
    ),
    stat = "identity",
    width = 0.99,
    alpha = 0.5
  ) +
  geom_text(
    aes(
      x = Sample,
      y = perc,
      label = paste0(round(perc, 1), "%")
    ),
    size = 3,
    position = position_stack(vjust = 0.5)
  ) + theme_test(base_size = 12) +
  scale_color_manual(values = c('gray10', "gray70")) +
  scale_fill_manual(values = c('gray10', "gray70")) +
  labs(y = "Read Distribution (%)", x = "") +
  theme(legend.title = element_blank()) +
  theme(axis.text.x = element_text(
    angle = 90,
    hjust = 1,
    vjust = 0.5,
    size = 10
  )) +
  scale_y_discrete(expand = c(0, 0.5)) + scale_x_discrete(expand = c(0, 0))
```


## Memory 8, 16 and 32 Threads Respectively
```{r,fig.height=5, fig.width=13, warning=FALSE, message=FALSE}
benchmark_rez_melted_memory <- read.csv("benchmark_rez_memory.csv")

memory8threads <-
  benchmark_rez_melted_memory %>% filter(Threads == "8 Threads") %>%
  ggplot(aes(
    x = factor(Sample, levels = order_df),
    y = (Value) / 1000000,
    group = Run
  )) +
  geom_point(aes(
    color = factor(Run),
    shape = factor(Run),
    fill = factor(Run)
  ), size = 2.5) +
  scale_shape_manual(values = c(3, 17, 16)) +
  scale_color_manual(values = global_colors) +
  labs(y = "Memory (Gigabytes)", x = "") +
  theme_light(base_size = 12) + theme(legend.position = "none",  legend.title = element_blank()) +
  scale_y_continuous(limits = c(30, 40)) +
  theme(strip.background = element_rect(fill = "white", colour = "white")) +
  theme(
    strip.text = element_text(colour = 'black'),
    strip.text.x =  element_markdown(hjust = 0.5, size = 12)
  ) +
  theme(axis.text.x = element_text(
    angle = 90,
    hjust = 1,
    vjust = 0.5,
    size = 10
  ))

memory16threads <-
  benchmark_rez_melted_memory %>% filter(Threads == "16 Threads") %>%
  ggplot(aes(
    x = factor(Sample, levels = order_df),
    y = (Value) / 1000000,
    group = Run
  )) +
  geom_point(aes(
    color = factor(Run),
    shape = factor(Run),
    fill = factor(Run)
  ), size = 2.5) +
  scale_shape_manual(values = c(3, 17, 16)) +
  scale_color_manual(values = global_colors) +
  labs(y = "", x = "") +
  theme_light(base_size = 12) + theme(legend.title = element_blank()) +
  scale_y_continuous(limits = c(30, 40)) +
  theme(strip.background = element_rect(fill = "white", colour = "white")) +
  theme(
    strip.text = element_text(colour = 'black'),
    strip.text.x =  element_markdown(hjust = 0.5, size = 12),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 10
    ),
    legend.position = "none"
  )

memory32threads <-
  benchmark_rez_melted_memory %>% filter(Threads == "32 Threads") %>%
  ggplot(aes(
    x = factor(Sample, levels = order_df),
    y = (Value) / 1000000,
    group = Run
  )) +
  geom_point(aes(
    color = factor(Run),
    shape = factor(Run),
    fill = factor(Run)
  ), size = 2.5) +
  scale_shape_manual(values = c(3, 17, 16)) +
  scale_color_manual(values = global_colors) +
  labs(y = "", x = "") +
  theme_light(base_size = 12) + theme(legend.title = element_blank()) +
  scale_y_continuous(limits = c(30, 40)) +
  theme(strip.background = element_rect(fill = "white", colour = "white")) +
  theme(
    strip.text = element_text(colour = 'black'),
    strip.text.x =  element_markdown(hjust = 0.5, size = 12),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  theme(axis.text.x = element_text(
    angle = 90,
    hjust = 1,
    vjust = 0.5,
    size = 10
  ))


plot_grid(
  memory8threads,
  memory16threads,
  memory32threads,
  nrow = 1,
  ncol = 3,
  labels = c("(a)", "(b)", "(c)"),
  rel_widths = c(0.85, 0.77, 1.115),
  hjust = 0.01,
  align = 'v'
)
```


## Ref_Bias_Per_Sample
```{r, warning=FALSE, message=FALSE}
df_combined <- read.csv("df_combined.csv")

(pe <-
    df_combined %>% filter(
      state1 == "Alignment Passed WASP Filtering"  |
        state1 == "Alignment Failed WASP Filtering"
    ) %>%
    ggplot(aes(
      x = Sample,
      y = diff,
      group = Sample,
      fill = state1
    )) +
    geom_col(alpha = 0.9) +
    scale_y_continuous(limits = c(-50, 100)) +
    theme_cowplot(font_size = 12) +
    geom_point(color = "gray20", size = 1) +
    scale_color_manual(
      values = c(
        "Alignment Passed WASP Filtering" = "skyblue2",
        "Alignment Failed WASP Filtering" = "steelblue4"
      )
    )  +
    scale_fill_manual(
      values = c(
        "Alignment Passed WASP Filtering" = "skyblue2",
        "Alignment Failed WASP Filtering" = "steelblue4"
      )
    )  +
    labs(
      y = paste0("Reference Bias ", "\n", "(REF% - ALT%)"),
      x = ""
    ) +
    geom_vline(xintercept = 0) +
    geom_hline(
      yintercept = 0,
      color = "gray40",
      linetype = "dashed",
      linewidth = 0.3
    ) +
    theme(
      legend.position = "right",
      legend.text = element_text(size = 8) ,
      legend.title = element_blank(),
      axis.text.x = element_text(
        angle = 90,
        hjust = 1,
        vjust = 0.5,
        size = 10
      )
    )
)
```


```{r, fig.height=11, fig.width=9, warning=FALSE, message=FALSE}
plot_df <- read.csv("plot_df.csv")
(b <- plot_df %>%
    filter(df_state == "vw_Tag1" | df_state == "vw_Tag_other") %>%
    ggplot(aes(
      x = reorder(pass_fail,-freq), freq,
      fill = Bias_State
    )) +
    geom_bar(
      stat = "identity",
      width = 0.99,
      alpha = 0.7
    ) +
    geom_text(
      aes(x = pass_fail,
          y = freq,
          label = freq),
      size = 4,
      position = position_stack(vjust = 0.5)
    ) +
    scale_fill_manual(values = c("gray45", "gray65")) +
    theme_light(base_size = 13) + ggtitle("") +
    theme(legend.position = "right", legend.title = element_blank()) +
    labs(x = "", y = "Percentage of Reads Aligning to REF/ALT") +  scale_x_discrete(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) + theme(plot.title = element_text(hjust = 0.5)) +
    facet_wrap( ~ Sample) + theme(strip.background = element_rect(
      fill = "white", colour = "white"
    )) +
    theme(
      strip.text = element_text(colour = 'black'),
      strip.text.x = element_markdown(hjust = 0.5)
    ) +  theme(axis.text.x = element_text(
      angle = 90,
      hjust = 0.5,
      vjust = 0.5,
      size = 10
    ))
)
```

## Mapping speed for STAR, WASP and STAR+WASP - Shared Computing Environment
```{r, fig.height=5, fig.width=12, warning=FALSE, message=FALSE}
merged_df <- read.csv("merged_df_speed_Elzar.csv")

merged_df$Threads <-
  ordered(merged_df$Threads ,
          levels = c("8threads", "16threads", "32threads"))

pp <- merged_df %>%
  ggplot(aes(
    x = factor(Sample, levels = order_df),
    y = reads_per_hour / 1000000,
    group = Run,
    color = Run
  )) +
  geom_point(aes(
    color = factor(Run),
    shape = factor(Run),
    fill = factor(Run)
  ), size = 2) +
  scale_shape_manual(values = c(3, 17, 16)) +
  facet_wrap( ~ Threads) +
  scale_color_manual(values = global_colors) +
  labs(y = paste0("Average Speed Across Runs", "\n", "(Million Reads/Hour)"),
       x = "") +
  theme_light(base_size = 12) + theme(legend.title = element_blank(), legend.position = "right") +
  theme(strip.background = element_rect(fill = "white", colour = "white")) +
  theme(
    strip.text = element_text(colour = 'white'),
    strip.text.x =  element_markdown(hjust = 0.5, size = 12)
  ) +
  theme(axis.text.x = element_text(
    angle = 90,
    hjust = 1,
    vjust = 0.5,
    size = 10
  )) +
  scale_y_log10(
    breaks = c(
      0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000
    ),
    limits = c(10, 1000),
    labels = c(
      "", "10", "", "", "", "", "", "", "", "", "100", "", "", "", "", "", "", "", "", "1000"
    )
  ) +
  annotation_logticks(
    sides = "l",
    outside = F,
    short = unit(0.05, "cm"),
    mid = unit(0.05, "cm"),
    long = unit(0.2, "cm")
  ) +
  coord_cartesian(clip = "off")

pp + geom_text(
  data = merged_df,
  mapping = aes(x = 4,
                y = 999,
                label = Threads),
  col = "gray30",
  fontsize = 12,
  fontface = "italic"
)
```

## Memory - STAR, WASP and STAR+WASP - Shared Computing Environment
```{r, fig.height=5, fig.width=6, warning=FALSE, message=FALSE}
merged_df <- read.csv("merged_df.csv")

unique(merged_df$Threads)

merged_df$Threads <-
  ordered(merged_df$Threads ,
          levels = c("8threads", "16threads", "32threads"))

merged_df %>%
  ggplot(aes(
    x = factor(Sample, levels = order_df),
    y = average_maxrss,
    group = Run
  )) +
  geom_point(aes(color = factor(Threads), shape = factor(Run), ), size =
               3) +
  scale_shape_manual(values = c(
    "STAR" = 3,
    "WASP" = 1,
    "STAR_WASP" = 17
  )) +
  scale_color_manual(values = c("orange2", "dodgerblue4", "firebrick4")) +
  labs(
    y = paste0(
      "Average Memory Across Runs",
      "\n",
      "(Maximum Resident Set Size (GB))"
    ),
    x = ""
  ) +
  theme_bw() + theme(legend.title = element_blank()) +
  theme(strip.background = element_rect(fill = "white", colour = "white")) +
  theme(strip.text = element_text(colour = 'black'),
        strip.text.x = element_markdown(hjust = 0)) +
  theme(axis.text.x = element_text(
    angle = 90,
    hjust = 1,
    vjust = 0.5,
    size = 10
  )) 
```





